---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "VizMasters - Team member 1, Team member 2,..."
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(janitor)
```

# Project Overview

The objective of this project is to analyze cancer-related data obtained from the source <https://ourworldindata.org/cancer>. Cancer is a complex disease with significant public health implications, and understanding its epidemiology, trends, and impact is crucial for effective prevention, diagnosis, and treatment strategies.

# Reasons for Choosing the Dataset

-   The datasets sourced from <https://ourworldindata.org/cancer> offer comprehensive and reliable information on various aspects of cancer, including mortality rates, types, demographics, and geographic variations.

-   Analyzing these datasets can provide valuable insights into the prevalence, distribution, and risk factors associated with different types of cancer.

-   The availability of multiple datasets allows for a holistic analysis of cancer-related trends and patterns, contributing to a better understanding of the disease's global burden.

# Datasets:

The project utilizes the following CSV files obtained from the provided source, each containing information related to cancer statistics. Below is an overview of the data files:

**01_annual-number-of-deaths-by-cause:** This file contains information on the annual number of deaths attributed to various causes, including cancer and encompasses data from 6,840 records across 35 different health conditions, including neoplasms. It likely includes data on the total number of deaths per year due to cancer, as well as deaths attributed to specific cancer types.

**02_total-cancer-deaths-by-type:** With 6841 records, This file provides details on the total number of cancer-related deaths categorized by cancer type. It may include information such as the number of deaths caused by lung cancer, breast cancer, prostate cancer, etc.

**04_cancer-death-rates-by-age:** Datasets, with 5,472 records, This dataset presents cancer death rates categorized by age groups. It offers insights into how cancer mortality rates vary across different age demographics.

**05_share-of-population-with-cancer-crude**: This file contains data on the crude prevalence of cancer within the population. It likely includes the percentage or proportion of the population diagnosed with cancer, without adjusting for age or other factors.

**06_share-of-population-with-cancer-types:** This dataset provides information on the prevalence of different cancer types within the population. It may include the percentage of individuals diagnosed with specific cancer types such as lung cancer, colorectal cancer, etc.

**07_number-of-people-with-cancer-by-age:** This file contains data on the number of individuals diagnosed with cancer categorized by age groups. It offers insights into the distribution of cancer cases across different age demographics.

**11_cancer-deaths-rate-and-age-standardized-rate-index:** This dataset likely includes information on cancer death rates and age-standardized rate indexes. It may provide insights into trends in cancer mortality rates over time and across different regions, adjusted for age variations in the population.

```{r}
#| label: load-dataset
#| message: false
```

Chosen for their depth and breadth, these datasets collectively provide a detailed picture of cancer's impact globally. They serve as a valuable resource for public health analysis, informing policy-making, and guiding research into effective cancer prevention and treatment strategies. The data likely originates from reputable health data collection entities, including national health ministries and global health organizations, reflecting a standardized and authoritative approach to health data compilation across countries and years.

# Data Wrangling and Displaying the Data

```{r}
#| label: Data wrangling
#| message: false

# Loading the datasets

annual_deaths_by_cause <- read_csv("./data/01_annual-number-of-deaths-by-cause.csv")
cancer_deaths_by_type <- read_csv("./data/02_total-cancer-deaths-by-type.csv")
cancer_deaths_by_age <- read_csv("./data/04_cancer-death-rates-by-age.csv")
population_with_cancer_crude <- read_csv("./data/05_share-of-population-with-cancer-crude.csv")
population_with_cancer_types <- read_csv("./data/06_share-of-population-with-cancer-types.csv")
population_with_cancer_by_age <- read_csv("./data/07_number-of-people-with-cancer-by-age.csv")
standardized_death_age_rate_index <- read_csv("./data/11_cancer-deaths-rate-and-age-standardized-rate-index.csv")

# Cleaning them using janitor package

annual_deaths_by_cause <- annual_deaths_by_cause |>
  clean_names()
cancer_deaths_by_type <- cancer_deaths_by_type |>
  clean_names()
cancer_deaths_by_age <- cancer_deaths_by_age |>
  clean_names()
population_with_cancer_crude <- population_with_cancer_crude |>
  clean_names()
population_with_cancer_types <- population_with_cancer_types |>
  clean_names()
population_with_cancer_by_age <- population_with_cancer_by_age |>
  clean_names()
standardized_death_age_rate_index <- standardized_death_age_rate_index |>
  clean_names()


# Displaying the datasets

head(annual_deaths_by_cause)
head(cancer_deaths_by_type)
head(cancer_deaths_by_age)
head(population_with_cancer_crude)
head(population_with_cancer_types)
head(population_with_cancer_by_age)
head(standardized_death_age_rate_index)
```

# Questions

The two questions you want to answer.

# Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
