---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "VizMasters - Team member 1, Team member 2,..."
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(janitor)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

The dataset collection provided offers an expansive overview of cancer-related statistics, encapsulating aspects such as incidence, mortality, prevalence across different age groups, and types of cancer. This comprehensive ensemble includes six distinct datasets, each contributing unique insights into the global landscape of cancer.

The first dataset, "Annual Number of Deaths by Cause," encompasses data from 6,840 records across 35 different health conditions, including neoplasms. It spans various countries and regions, offering a broad perspective on global health trends and mortality, thereby serving as a fundamental resource for understanding the impact of cancer in relation to other major health conditions.

Focusing more closely on cancer-specific outcomes, the "Cancer Deaths by Age" and "Cancer Death Rates by Age" datasets, with 5,472 and 6,187 records respectively, break down mortality figures into different age brackets. This distinction is crucial for analyzing how cancer affects various age groups, with the former providing absolute numbers of deaths and the latter offering death rates per 100,000 people, allowing for standardized comparisons across populations.

The "Share of Population with Cancer (Crude)" dataset, through its 6,157 records, reveals the unadjusted proportion of populations diagnosed with cancer, highlighting the general burden of the disease on societies. In contrast, the "Share of Population with Cancer Types" dataset delves into the prevalence of specific cancer types (such as breast, prostate, and lung cancer) within the population, comprising 1,461 records. This granularity aids in identifying which cancers are most prevalent and requires targeted research and intervention strategies.

Lastly, the "Share of Population with Cancer by Age" dataset, with its 6,780 records, sheds light on cancer prevalence across various age demographics. This is instrumental in understanding the increased risk associated with aging and identifying which age groups are most vulnerable.

Chosen for their depth and breadth, these datasets collectively provide a detailed picture of cancer's impact globally. They serve as a valuable resource for public health analysis, informing policy-making, and guiding research into effective cancer prevention and treatment strategies. The data likely originates from reputable health data collection entities, including national health ministries and global health organizations, reflecting a standardized and authoritative approach to health data compilation across countries and years.

## Data Wrangling and Displaying the Data

```{r}
#| label: Data wrangling
#| message: false

# Loading the datasets

annual_deaths_by_cause <- read_csv("./data/01_annual-number-of-deaths-by-cause.csv")
cancer_deaths_by_type <- read_csv("./data/02_total-cancer-deaths-by-type.csv")
cancer_deaths_by_age <- read_csv("./data/04_cancer-death-rates-by-age.csv")
population_with_cancer_crude <- read_csv("./data/05_share-of-population-with-cancer-crude.csv")
population_with_cancer_types <- read_csv("./data/06_share-of-population-with-cancer-types.csv")
population_with_cancer_by_age <- read_csv("./data/07_number-of-people-with-cancer-by-age.csv")
standardized_death_age_rate_index <- read_csv("./data/11_cancer-deaths-rate-and-age-standardized-rate-index.csv")

# Cleaning them using janitor package

annual_deaths_by_cause <- annual_deaths_by_cause |>
  clean_names()
cancer_deaths_by_type <- cancer_deaths_by_type |>
  clean_names()
cancer_deaths_by_age <- cancer_deaths_by_age |>
  clean_names()
population_with_cancer_crude <- population_with_cancer_crude |>
  clean_names()
population_with_cancer_types <- population_with_cancer_types |>
  clean_names()
population_with_cancer_by_age <- population_with_cancer_by_age |>
  clean_names()
standardized_death_age_rate_index <- standardized_death_age_rate_index |>
  clean_names()


# Displaying the datasets

head(annual_deaths_by_cause)
head(cancer_deaths_by_type)
head(cancer_deaths_by_age)
head(population_with_cancer_crude)
head(population_with_cancer_types)
head(population_with_cancer_by_age)
head(standardized_death_age_rate_index)
```

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
